---
title: "R Notebook"
output: 
        html_notebook:
                theme: journal
---

#Statistics Examples

##Linear Regression
*http://r-statistics.co/Linear-Regression.html*


```{r}
head(cars)
```

### Scatter Plot
- Visualize the linear relationship between the predictor and the response.

```{r}
scatter.smooth(x=cars$speed, y=cars$dist, main="Dist - Speed")
```

###BoxPlot - Check for outliers
- Any datapoint outside the 1.5* interquartile range is considered an outlier. IQR is the distance between the 25th and 75th percentiles for that variable.

```{r}
par(mfrow=c(1,2)) #divide graph area into 2 columns
boxplot(cars$speed, main = "Speed", sub = paste("Outlier rows: ", boxplot.stats(cars$speed)$out)) #box plot for speed
boxplot(cars$dist, main="Distance", sub = paste("Outlier rows: ", boxplot.stats(cars$dist)$out))  #box plot for distance
```

### Density Plot - Check if the response variable is close to normality

```{r}
library(e1071)
par(mfrow = c(1, 2))  # divide graph area in 2 columns
plot(density(cars$speed), main = "Density Plot: Speed", ylab = "Frequency", sub = paste("Skewness: ", round(e1071::skewness(cars$speed), 2)))   # density plot for speed
polygon(density(cars$speed), col = "red")
plot(density(cars$dist), main = "Density Plot: Distance", ylab = "Frequency", sub = paste("Skewness:", round(e1071::skewness(cars$dist), 2)))  # density plot for distance
polygon(density(cars$dist), col = "red")
```

### Correlation

- Defined as the statistical measure of the linear dependence of two variables. Can take values between -1 and 1.
Closer to 1 indicates a strong positive correlation. An inverse relationship shows values closer to -1. 
- Values close to 0 indicate weak relationship; should look for better descriptors.

```{r}
cor(cars$speed, cars$dist)   # calculate correlation between speed and distance
```

###Build Linear Model
` lm() ` is the function for building linear models.

```{r}
linearMod <- lm(dist ~ speed, data = cars)  # build linear regression model on full data
print(linearMod)
```

**dist = Intercept + (b * speed)**

=> dist = -17.579 + 3.932 * speed

### Linear Regression Diagnostics

**MUST check for statisical significance before using model.**

print summary stats for the linear model:
```{r}
summary(linearMod)  # model summary
```


